---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# growr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

growr helps you with the pieces of growth curve analysis. 

### High Level Interface

Currently, growr has high level interface that is meant to be simple with reasonable defaults:

1. `preprocess()` transform raw measures data into values suitable for summarization.
2. `summarise_fit()` summarise a fit method into growth-curve metrics such as the fit's $A$, $\mu$, $\lambda$.
3. `add_grouping()` and `add_groupings()` provide a convenient way to add relationships between sets of wells of a microtitre plate. Reference and target well designations can then be used to rescale summary metrics within groups (with for example, the provided `relatively()`).

### lower level functions

growr also exports an extensive set of models that make up a low level interface. Nearly all these functions have previously been used in growth curve analyses. *The majority of this functionality is not new*. It reflects a collection from various public source from this author's exploration. Citations provided in documentation of the individual methods.

## Example


#### Preprocessing

`preprocess()` takes a vector of values and returns a transformed version. The details of the transformation depend on the argument choices but follows the preprocessing steps laid out in the precog paper.

```{r preprocess}
library(tidyverse)
library(growr)

data <- mtpview::mtp_example3
data
data %>% 
  group_by(well) %>% 
  mutate(measure_pp = preprocess(measure))

```

#### Summarising fit

```{r}
well_summaries <- data %>% 
  group_by(well) %>% 
  mutate(measure_pp = preprocess(measure)) %>% 
  summarise(metrics = list(summarise_fit(x = runtime, y = measure_pp))) %>% 
  unnest(cols = c(metrics))
```


#### Grouping

`add_groupings()` takes one or more strings with grouping shorthand, expands them and joins them onto the provided data. The `source well(s)`->`target well(s)`. Sets of wells are specified with the familiar rectanuglar colon style notation: `A01:A12` is expanded to contain row A wells from column 1 to column 12 inclusive. 

```{r groups}
well_summaries %>% 
  add_groupings(c("A01->A02:A12", "B01:H01->B02:H12"), nrow = 8, ncol = 12)
```

A more complex example (using the internal fxn for demonstration of the parsing that goes on under the hood of `add_groupings()`)

```{r}
growr:::parse_grouping('p1!A01,B01:B04,H03->A01:H12', 8 , 12)
```

#### Relatively

Rescale you data by group, for example:

```{r}
well_summaries %>% 
  add_groupings(c("A01->A02:A12", "B01:H01->B02:H12"), nrow = 8, ncol = 12) %>% 
  group_by(group) %>% 
  mutate(relative_mu = relatively(mu, is_ref)) 
```

---

## Installation

``` r
# install.packages("growr")
remotes::install_github("npjc/growr")
```

---

<!-- ### Other examples -->


<!-- ```{r preprocess} -->
<!-- library(tidyverse) -->
<!-- library(mtpview) -->
<!-- library(growr) -->

<!-- data <- mtpview::mtp_example3 -->
<!-- one_well <- data %>% filter(well == 'A01') -->

<!-- one_well %>% -->
<!--     mutate(measure_pp = preprocess(measure, log_base = F, bg_subtract = F)) %>% -->
<!--     ggplot(aes(x = runtime)) + -->
<!--     geom_point(aes(y = measure)) + -->
<!--     geom_point(aes(y = measure_pp), size = 0.25, color = "red") + -->
<!--     theme_light() + -->
<!--     theme(axis.ticks.y = element_blank(), -->
<!--           axis.ticks.length = unit(.pt * 2, "pt"), -->
<!--           panel.grid.minor.y = element_blank(), -->
<!--           panel.grid.major.x = element_line(linetype = 'dashed', color = 'grey80'), -->
<!--           panel.grid.minor.x = element_line(linetype = 'dashed', color = 'grey90'), -->
<!--           axis.text.x = element_text(hjust = 0, vjust = 1, margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))) -->

<!-- ``` -->

