---
title: "Implementing Gompertz and recipes"
author: "Nicolas Coutin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load}
# library(tidyverse)
# library(recipes)
```

what is the model fitting I'm trying to do?

1. primary model : od as fxn of time


2. secondary model: the derived values $\mu$ $\lambda$ $lag$


## Primary Models

Gompertz's standard equation:

...parametrized for $A$, $\mu$ and $\lambda$

$$
y(t) = A \cdot exp [-exp(\frac{\mu \cdot e}{A}(\lambda - t) + 1)]
$$

Start from the a,b,c form:

$$
y = a \cdot exp[-exp(b -cx)]
$$

Maximum growth rate is defined as second derivative going to 0 (inflection point):

$$
\frac{d^2y}{dt^2} = 
a c \cdot exp(-exp(b - c t) + b - c t) \cdot (c \cdot exp(b - c t) - c)
$$
alternate form:

$$
\frac{d^2y}{dt^2} = 
a c^2 \ exp[-exp(b - c t) + b - c t] \cdot [exp(b - c t) - 1]
$$
root for t is given by

$$
t_i = \frac{b}{c}
$$

now substitute that result into first derivative:

$$
ac \cdot exp[-exp(b - c \cdot t_i)] \cdot exp(b - c \cdot t_i)
$$

$$
ac \cdot exp[-exp(b - c \cdot \frac{b}{c})] \cdot exp(b - c \cdot \frac{b}{c})
$$

$$
ac \cdot exp[-exp(b - b)] \cdot exp(b - b)
$$
$$
ac \cdot exp[-exp(0)] \cdot exp(0)
$$
$$
ac \cdot exp[-1] \cdot 1
$$

$$
\mu_m = ac \cdot exp[-1] = 0.3679ac = \frac{ac}{exp(1)} = \frac{ac}{e}
$$
So c in grompertz equation can be substituted for:

$$
c = \frac{\mu_m \cdot e}{a}
$$

tangent line throught the inflection point:

$$
y = \mu_m \cdot t + b
$$
$$
y = \mu_m \cdot t + b
\\
\frac{y - b}{t} = \mu_m = 
$$

$$
y = a \cdot exp[-(exp(b) / exp(cx)]
$$


$$
y(t) = \alpha \cdot exp(-\beta \cdot exp(-k^t))
$$


As an `R` function:

```{r}
#' The function calculates the values of the Gompertz growth curve for given time points.
#'
#' @param time Time points (x-axes) for which the function values will be returned.
#' @param A Maximum of the curve.
#' @param mu Maximum slope.
#' @param lambda Lag phase
grofit_gompertz <- function (time, A, mu, lambda) {
    A * exp(-exp(mu * exp(1) * (lambda - time) / A + 1.0))
}
```

use this to generate some data with known values:

- 96 time points, one every 15 mintues.
- maximum of 2
- growth rate of 0.005

```{r}
time <- 1:96 * 15
y <- grofit_gompertz(time, A = 2, mu = 0.005, lambda = max(time) * 0.1) + rnorm(96,mean = 0.01, sd = 0.05)
data <- data.frame(time, y)
plot(data)
```


Formula:

```r
y ~ gompertz(x, A, mu, lambda)
```

```{r}
data <- data$y
time <- time
formula <- y ~ grofit_gompertz(time, A, mu, lambda)
fit <- nls(formula, 
           data = data,
           start = list(A = 1, mu = 0.01, lambda = 120))
plot(fit)
coef(fit)
coef(summary(fit))
broom::tidy(fit)
broom::glance(fit)
broom::augment(fit) %>% 
  ggplot(aes(x = time)) + 
  geom_point(aes(y = y)) +
  geom_line(aes(y = .fitted), color = 'blue')

broom::augment(fit) %>% 
  ggplot(aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

```

